# Practical 4

We'll start by loading the necessary packages.

```{r load libraries, echo = TRUE, message = FALSE, warning = FALSE}
library("sf")
library("dplyr")
library("tmap")
library("jrSpatial")
```

Question 1. # THEO We could make them create a geometry point instead with `st_point()`?

1. Create a dataframe representing our current location and it's longitude and latitude. `loc = data.frame(lon = , lat = )`. You can use Google Maps to find our current longitude and latitude.

```{r create lon lat}
loc = data.frame(lon = -1.6053, lat = 54.9705)
```

2. In practical 1, we learned how to create a **sf** data frame from scratch using `st_sf()`. Because our coordinates are just numeric columns instead of a single geometry column, we can't use `st_sf()`. Instead we have to use `st_as_sf()` and specify the coordinate columns using the `coords` argument, like so 
```{r sf convert, echo = TRUE}
loc = st_as_sf(loc, coords = c("lon", "lat"))
```

3. Inspect the dataframe. Use the function `st_is_longlat()` to check whether the data is recognised as using longlat degrees. What about `st_crs()`?

```{r longlat check}
st_is_longlat(loc)
st_crs(loc)
# no longlat because we have not set the coordinate ref system yet
```

4. Set the CRS to have epsg 4326 using `st_set_crs()` and then check `st_is_longlat()` again. 

```{r set crs longlat check}
loc = st_set_crs(loc, 4236)
st_is_longlat(loc)
```



Question 2. 

1. Load the shape files for New Zealand using `data(nz, package = "jrSpatial")`. Plot the data using `tm_shape()` with `tm_borders()`. What do you notice?

```{r nz}
data(nz, package = "jrSpatial")
tm_shape(nz) + tm_borders() + tm_fill()

# Part of New Zealand is missing. 
```

2. The Canterbury region of New Zealand is missing from the data. Load in the Canterbury data using 
```{r, echo = TRUE}
data(canterbury, package = "jrSpatial")
```
Say we have two spatial data frame, `x` and `y`, we can row bind them by using 
```{r, echo = TRUE, eval = FALSE}
rbind(x, y)
```
Try to use `rbind()` to add the Canterbury row to the original New Zealand data. What does the output tell you?

```{r cant bind, eval = FALSE}
nz = rbind(nz, canterbury)
# The two data sets have different CRS. R won't let us combine them.
```

3. The two data sets have different CRS. Find out the CRS of the New Zealand data using `st_crs().` Then use `st_transform()` to transform the Canterbury data to use the same CRS as the rest of New Zealand. Combine the data with `rbind()` and plot the output. 

```{r transform}
st_crs(nz)
canterbury = st_transform(canterbury, 4326)

nz = rbind(nz, canterbury)

tm_shape(nz) + tm_borders() + tm_fill()
```



