This practical is all about Methamphetamines! Yes, you read that right. Let's load in the data and necessary packages first
```{r, message = FALSE, warning = FALSE}
library("sf")
library("leaflet")
library("dplyr")
data(meth, package = "jrSpatial")
```

This is a dataset containing locations of real meth lab busts in the US. Take a look at the data to get a feel of it
```{r}
head(meth)
```
Confession: I've made up the pounds_seized column. Unfortunately the DEA don't give that kind of stuff out. However I've checked a couple of records and the numbers are within reason! 

1. Using **leaflet**, create an interactive map of the locations of the meth lab busts. Hint: Use `addCircles()`
```{r, fig.keep="none"}
leaflet(meth) %>% 
  addTiles() %>% 
  addCircles(data = meth)
```

2. Try adding the argument `radius = ~pounds_seized` to `addCircles()`. What happens?
```{r, fig.keep="none"}
leaflet(meth) %>% 
  addTiles() %>% 
  addCircles(data = meth, 
             radius = ~pounds_seized)
# radius of circles is bigger if more pounds of meth were seized
```

3. Try colouring the points by the pounds seized. To do this, use the `colourNumeric()`
```{r, fig.keep="none"}
pal_size = colorNumeric(palette = c("green", "red"), 
                        domain = meth$pounds_seized)

leaflet(meth) %>% 
  addTiles() %>% 
  addCircles(data = meth, 
             radius = ~pounds_seized, 
             color = ~pal_size(pounds_seized))
```

4. The problem with using `colourNumeric()` in this instance, is that the data is heavily skewed. If we look at a histogram of the pounds_seized, there isn't many meth labs that had over 1000 pounds of meth seized
```{r, echo = TRUE, results="show"}
library("ggplot2")
ggplot(meth, aes(x = pounds_seized)) + 
  geom_histogram()
```

5. We can set up a colour scheme such that we have discrete colours using `colourBin()`, like so
```{r, echo = TRUE}
pal_size_bin = colorBin(palette = c("Green", "Yellow", "Orange", "Red"), 
                        domain = meth$pounds_seized, 
                        bins = c(0,150,300,750,max(meth$pounds_seized)))
```
What colour scheme do you think we're using here?
```{r}
# green - from 0 -150
# yellow - from 150 - 300
# orange - from 300 - 750
# red - from 750 - max value in pounds_seized
```

6. Use the new colour scheme to colour the points
```{r, fig.keep="none"}
leaflet(meth) %>% 
  addTiles() %>% 
  addCircles(data = meth, 
             radius = ~pounds_seized, 
             color = ~pal_size_bin(pounds_seized))
```

7. Add a legend to the plot
```{r, fig.keep="none"}
leaflet(meth) %>% 
  addTiles() %>% 
  addCircles(data = meth, 
             radius = ~pounds_seized, 
             color = ~pal_size_bin(pounds_seized)) %>% 
  addLegend(title = "Pounds seized",
            pal = pal_size_bin, 
            values = ~pounds_seized)
```
